FINE SCALE REGISTRATION OF SIDEWALKS AND OTHER RIBBON-LIKE FEATURES FROM AERIAL IMAGERY

### 2018-07-27 Chapters/3.Hypothesis/Hypothesis.tex
Fixing format and images

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.95\columnwidth]{Figures/ribbon-3d-combined.pdf}
    \caption[3D Ribbon Image]{An illustration of the parameters for a ribbon: (top) the $k$ axis captures the thickness of the ribbon and the $i$ axis captures the distance along the initial ribbon, the $j$ axis records horizontal displacements from the initial ribbon's medial curve; (bottom) the boundaries of the ribbon must lie within $\pm \MaxDistance{}$, pixels within $\pm \MinRadius{}$ of the center are probably part of the ribbon and pixels further than $\pm \MaxRadius{}$ are likely negative. }
    \label{fig:ribbon_3d}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{Figures/sw-figure.pdf}
    \caption[2D Ribbon Image]{A demonstration of the parameters for a ribbon in 2D. Where $k_n$ indicates a parameter that describes the radius of the medial curve that we considered to be part of the ribbon, $k_f$ is a distance from the medial curve that we considered very unlikely to be part of the ribbon, and $k_W$ is the maximum width of the ribbon image.}
    \label{fig:2d_ribbon}
\end{figure}

As shown in \ref{fig:2d_ribbon}, we assume that there is some uncertainty in the position $\mathbf{x}_s$ of the medial curve and also uncertainty in the radius $r_s$ of the discredited ribbon, but that there is range of distances $t \in 0\dots k_n$ from the medial curve which we can accept labeling as positive examples that are within the ribbon, and that there is some range of distances $t \in k_f\dots k_W$ which we can accept labeling as negative sample, and finally a range $t\in (k_n+1)...(k_f-1)$ of distances that are possible locations of the ribbon's boundary.  Therefore, when searching for improved medial curves and radii of the ribbon, we can consider a discrete range of offsets $t=-k_W,..., k_W$ chosen to be wide enough that true ribbon image itself likely fits entirely within the grid of input negative samples along its edges.

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/GMM_needed.png}
    \caption[\ac{GMM} Result 1]{Example of GMMs on sample sidewalk. The first row and third row shows the foreground GMM for sample sidewalk, and the second row and fourth row shows the back-ground for simple sidewalk. We use grey-scale to show the changes of probabilities, where for foreground, white is more likely to be sidewalk, and black is not.}
    \label{fig:GMM_result}
\end{figure}

>>>>

\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{Figures/ribbon-3d-combined.pdf}
    \caption[3D Ribbon Image]{An illustration of the parameters for a ribbon: (top) the $k$ axis captures the thickness of the ribbon and the $i$ axis captures the distance along the initial ribbon, the $j$ axis records horizontal displacements from the initial ribbon's medial curve; (bottom) the boundaries of the ribbon must lie within $\pm \MaxDistance{}$, pixels within $\pm \MinRadius{}$ of the center are probably part of the ribbon and pixels further than $\pm \MaxRadius{}$ are likely negative. }
    \label{fig:ribbon_3d}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{Figures/sw-figure.pdf}
    \caption[2D Ribbon Image]{A demonstration of the parameters for a ribbon in 2D. $k_n$ indicates a parameter that describes the radius of the medial curve that we considered to be part of the ribbon, $k_f$ is a distance from the medial curve that we considered very unlikely to be part of the ribbon, and $k_W$ is the maximum width of the ribbon image.}
    \label{fig:2d_ribbon}
\end{figure}

As shown in \ref{fig:2d_ribbon}, from figure \ref{fig:osm_oxford_path}, we should assume that there is some uncertainty in the position $\mathbf{x}_s$ of the medial curve and also uncertainty in the radius $r_s$ of the discredited ribbon. 
But that there is range of distances $t \in 0\dots k_n$ from the medial curve which we can accept labeling as positive examples that are within the ribbon, and that there is some range of distances $t \in k_f\dots k_W$ which we can accept labeling as negative sample. 
Finally a range $t\in (k_n+1)...(k_f-1)$ of distances that are possible locations of the ribbon's boundary.
Therefore, when searching for improved medial curves and radii of the ribbon, we can consider a discrete range of offsets $t=-k_W,..., k_W$ chosen to be wide enough that true ribbon image itself likely fits entirely within the grid of input negative samples along its edges.

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/gmm_sample_1.png}
    \caption[\ac{GMM} Result 1]{Example of GMMs on sample sidewalk. The first row and third row shows the foreground GMM for sample sidewalk, and the second row and fourth row shows the back-ground for simple sidewalk. We use grey-scale to show the changes of probabilities, where for foreground, white is more likely to be sidewalk, and black is not.}
    \label{fig:GMM_result}
\end{figure}

### 2018-07-27 Chapters/3.Hypothesis/Paper.tex
Fixing format and images

\chapter{Publication}

@@

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.9\textwidth]{Figures/diagram.png}
\caption[Framework Overview 2]{Our approach adopted to predict precise boundaries for ribbon-like features. Where black line indicated a sidewalk's geometric information, and red lines shown our result for the sidewalk boundaries. From (a) to (f), the 6 step approach introduce input map (a), initial trajectory (b), ribbon image generation (c), density Estimation for pixels(d), apply \ac{DP} solution(e), reshape sidewalk(f).}
\label{fig:Apparatus}
\end{center}
\end{figure}

\section{Warping an input image}

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{Figures/Sample4_needed.png}
    \caption[Sample Sidewalk 1]{A ribbon image straighten output which used the previous process from figure \ref{fig:StraightenProcess} to convert an original image into ribbon image .}
    \label{fig:Sample_Sidewalk_4}
\end{figure}

>>>>

\chapter{Experiment}

@@

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.8\textwidth]{Figures/diagram6.png}
\caption[Framework Overview]{Our approach adopted to predict precise boundaries for ribbon-like features. Where black line indicated a sidewalk's geometric information, and red lines shown our result for the sidewalk boundaries. From (a) to (f), the 6 step approach introduce input map (a), initial trajectory (b), ribbon image generation (c), density Estimation for pixels(d), the result from our \ac{DP} solution(e), the output result by reshape sidewalk(f).}
\label{fig:Apparatus}
\end{center}
\end{figure}

\section{Warping an Input Image}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/Sample4_needed.png}
    \caption[Sample Sidewalk 1]{A ribbon image straighten output which used the previous process from figure \ref{fig:StraightenProcess} to convert an original image into ribbon image .}
    \label{fig:Sample_Sidewalk_4}
\end{figure}

### 2018-07-27 Chapters/4.Expect_results/Results.tex
Fixing format and images

\chapter{Preliminary Result}

A four-step adopt approach for sidewalk segmentation was developed. Locating sidewalk geometric information from raw \ac{VHR} data would be the first step, then generating ribbon image on each sidewalk which has been locate in the map data. After that, adopting density estimation on each pixel in ribbon image which generated from last step. Finally, we applied dynamic programming algorithm with fully developed penalty function to get edges on both side of sidewalks, then just simply converted the ribbon image back to it's original shape.

\section{Sidewalk Geometric Information Locator}

Before using deep learning network to determine sidewalk's location, hand marking all sidewalk from imagery or extract the geometric information of each sidewalk from resources like \ac{OSM} was required. It's challenging since the geometric accuracy was not precise, which could cause performance issue due to the fact that we relied on this information to segment the sidewalk part from maps for the next step. To ensure data accuracy, we spent days marking all sidewalks and their precise boundaries in Arizona Central. We can located each sidewalk by using the geometric information and saved them into separate files. We also calculated the average sidewalk width from boundaries to decide reasonable window size in the following step.

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{Figures/straghten.pdf}
    \caption[Ribbon Image Generation]{Ribbon image generating process: from sub-figures A to G, we showed the step by step process to generate ribbon-image from a partial sidewalk. From raw input (A), we smoothed mid line and generated parallel off-site (B). Generated pixels (C) and connected each pixel to its correspond color data from the raw input (D). After gathered pixel data (E), we removed mid-line (F) and generated the ribbon image(G).}
    \label{fig:StraightenProcess}
\end{figure}

@@

To apply the hypothesis, we develop a way to convert the given sidewalk into a ribbon image. As shown in figure \ref{fig:StraightenProcess}, first, we used the previous outcome (A) from the last step as input and connected all coordinate points data as the mid-line for the sample sidewalk, then we applied a smooth function that we created to smooth the mid-line, so the mid-line shape would fit better to the actual sidewalk shape(B). We generated two parallel line as offsets on each side of the mid-line with a distance of twice the average width in between. So ideally even the mid-line was not perfectly laying on the middle, we could still fit the whole sidewalk into the ribbon image. For each point on the mid-line, we connected the corresponding points on both offsets to get the movement of the sidewalk and recorded them(C, D), we could generate the straightened ribbon image using the color data from each pixel in the points set that we record earlier(E). The image data just in between each offset, with constant image width could be retrieved after that(F). Reshaping the image data was needed to generate the ribbon image(G). Figure \ref{fig:Sample_Sidewalk_2} shows the complete ribbon-image output result on sample sidewalk. We separate it into parts to show better detail.

@@

Before moving on to dynamic programming approach, each pixel need to be given a initial score to classify each pixel is sidewalk or not. To achieve that, as shows in figure \ref{fig:GMM_Sample_2}, we applied a density estimation function to find the probability for each pixel in the ribbon image. For foreground, we used grey-scale to determine the probability likely-hood for the sidewalk-feature, where white is more likely to be sidewalk and black is less likely. The background is the other way around, which made black is more likely to be sidewalk and white is not.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.65\textwidth]{Figures/GMM_SAMPLE2.png}
    \caption[Density Estimation on Sample Sidewalk]{A sample figure with density estimation applied on figure \ref{fig:Sample_Sidewalk_2}. Where white indicates pixels are more likely to be a sidewalk feature and black indicates pixels are more likely to be a non-sidewalk feature. We separate it into 4 parts from column 1 to 4.}
    \label{fig:GMM_Sample_2}
\end{figure}

@@

We used our dynamic programming algorithm on the ribbon image to find the max score for each pixel that had more probability to be sidewalk feature that we mentioned in Chapter 4. More importantly, we introduced a penalty control function that makes each horizontal movement in the edges or center cost more than vertical movement. Basically, for each line of pixels, if they think the edges or center should shift more horizontally than one pixel compares to the last line, we score them with more penalty.

Figure \ref{fig:penalty} shows our algorithm managed to separate the driveway and sidewalk under the same feature as penalty increase. So we could apply the dynamic programming solution to locate the precise boundaries of given ribbon image. The penalty control improved our performance significantly. Because of this, our approach could predict the sidewalk under the tree shadow and other shadows or obstacles, which bring us more accurate result compare to other methods.

As shown in figure \ref{fig:sample_result2}, after we generate the result on ribbon image, just need to simply covert the ribbon image back to its original shape by reverse the method that generates the ribbon image.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/penalty.pdf}
    \caption[Penalty Process]{Comparison with penalty change, we tested a variety of variables to decide which variable had best performance for width control. With the penalty increased at a reasonable amount, our algorithm decided not to recognize a driveway as sidewalk walk, in order to keep the sidewalk width consistent, even though they are the same texture.}
    \label{fig:penalty}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{Figures/sample2_result.png}
    \caption[Desire Output on Sample Sidewalk]{Detail demonstration on desire result after applying our dynamic programming approach. Where row 1 shows the result after reshape, row 2 to 5 shows the boundaries we calculated on ribbon image from \ref{fig:Sample_Sidewalk_2}.}
    \label{fig:sample_result2}
\end{figure}

@@

We also test our approach in different cities. For initial trajectories, We pull the corresponding footway data from overpass-turbo \cite{overpass_turbo}. Finding map data is challenging since there not exist much \ac{VHR} imagery from public source. We use the imagery data form GIS government database and treat it as input map. \figref{fig:oxford1} demonstrate plausible result on a sidewalks heavy area, we randomly selected a few from Oxford, OH. Row 1 shows that our approach are able to find the precise boundaries under trees or building shadows, with less accuracy around intersection area. Except above from row1, row 2 also demonstrate the camouflage with adjacent material. We treat the intersection part as overlapping sidewalks, and process them individually. \figref{fig:ny1}, \figref{fig:ny2} and \figref{fig:ny3} show the random samples we select from New York area. Predicting is heavier involved in these area since some sidewalks are partially under tree leaves instead of just shadows.

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{Figures/ny2.png}
    \caption[Sample Sidewalk 5]{Demonstration on sample sidewalks with our approach. We use sidewalks that are random selected from New York area. Row 1 shows a thicker sidewalk that camouflage with adjacent road and street. Row 2 shows complex sidewalk texture with irregular shape.}
    \label{fig:ny2}
\end{figure}

>>>>

\chapter{Preliminary Results}

We developed a six-step adopt approach for sidewalk segmentation. 
With the input \ac{VHR} imagery, we locate the sidewalk geometric information and it's initial trajectory, then we generate ribbon image on each sidewalk which has been locate in the map data. 
After that, we adopt density estimation on each pixel in the ribbon image. 
We apply the dynamic programming algorithm with fully developed penalty function to get precious boundaries on sidewalks, then just simply converted the ribbon image back to it's original shape.

\section{Validate Initial Trajectory}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/arizon_center.png}
    \caption[Demonstration on Arizona Center]{Detail demonstration on part of Arizona center, all sidewalks' \ac{GIS} information shows in Orange.}
    \label{fig:arizon_center}
\end{figure}

Before applying deep learning network to determine sidewalk's location, it's necessary to hand marking all sidewalk from imagery or extract the geometric information of each sidewalk from resources like \ac{OSM}. 
It's challenging since the geometric accuracy was not precise, which could cause performance issue since we rely on this information to segment the sidewalk part from maps for the next step. 
To ensure data accuracy, as shows in figure \ref{fig:arizon_center}, we spent days marking all sidewalks and their precise boundaries in Arizona Central for initial experiments. 
We can locate each sidewalk by using the geometric information and saved them into separate files. 
We also calculate the average sidewalk width from boundaries to decide reasonable window size in the following step.

@@

To apply the hypothesis, we develop a way to convert the given sidewalk into a ribbon image. 
As shown in figure \ref{fig:StraightenProcess}, we use the previous outcome (A) from the last step as input and connected all coordinate points data as the mid-line for the sample sidewalk. 
Then we apply a smooth function that we created to smooth the mid-line, so the shape of mid-line fits better to the actual sidewalk shape(B). 
We generate two parallel line as offsets on each side of the mid-line with a distance of twice the average width in between. 
Ideally even the mid-line is not perfectly laying on the middle, we would still fit the whole sidewalk into the ribbon image. 
For each point on the mid-line, we connect the corresponding points on both offsets to get the movement of the sidewalk and recorded them(C, D), we generate the straightened ribbon image by using the color data from each pixel in the points set that we record earlier(E). 
The image data in between each offset, with constant image width could be retrieved after that (F).
Then we reshape the image data to generate the ribbon image(G). 
Figure \ref{fig:Sample_Sidewalk_2} shows the complete ribbon-image output result on sample sidewalk. We separate it into parts to show better detail.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/straghten.pdf}
    \caption[Ribbon Image Generation]{Ribbon image generating process: from sub-figures A to G, we showed the step by step process to generate ribbon-image from a partial sidewalk. From raw input (A), we smoothed mid line and generated parallel off-site (B). Generated pixels (C) and connected each pixel to its correspond color data from the raw input (D). After gathered pixel data (E), we removed mid-line (F) and generated the ribbon image(G).}
    \label{fig:StraightenProcess}
\end{figure}

@@

@@

Before move on to dynamic programming algorithm, we need to give each pixels a initial score for classification, to classify each pixel is sidewalk or not. 
To achieve that, as shows in figure \ref{fig:GMM_Sample_2}, we applied a density estimation function to find the probability for each pixel in the ribbon image. 
For foreground density, we used grey-scale to determine the probability likely-hood for the sidewalk-feature, where white is more likely to be sidewalk and black is less likely. 
The background density is the opposite, which made black is more likely to be sidewalk and white is not. 
We use the log likely-hood function to calculate the foreground and background density and estimate density on pixels.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/GMM_SAMPLE2.png}
    \caption[Density Estimation on Sample Sidewalk]{A sample figure with density estimation applied on figure \ref{fig:Sample_Sidewalk_2}. Where white indicates pixels are more likely to be a sidewalk feature and black indicates pixels are more likely to be a non-sidewalk feature. We separate it into 4 parts from column 1 to 4.}
    \label{fig:GMM_Sample_2}
\end{figure}

@@

We use our dynamic programming algorithm on the ribbon image to find the max score for each pixel that had better probability to be sidewalk feature that we mentioned in Chapter 4. 
More importantly, we introduce a penalty control function that makes each horizontal movement in the edges or center cost more than vertical movement. Basically, for each line of pixels, if they think the edges or center should shift more horizontally than one pixel compares to the last line, we score them with more penalty.

Figure \ref{fig:penalty} shows our algorithm managed to separate the driveway and sidewalk under the same feature as penalty increase. So we could apply the dynamic programming solution to locate the precise boundaries of given ribbon image. The penalty control improved our performance significantly. Because of this, our approach could predict the sidewalk under the tree shadow and other shadows or obstacles, which bring us more accurate result compare to other methods.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/penalty.png}
    \caption[Penalty Process]{Comparison with penalty change, we tested a variety of variables to decide which variable had best performance for width control. With the penalty increased at a reasonable amount, our algorithm decided not to recognize a driveway as sidewalk walk, in order to keep the sidewalk width consistent, even though they are the same texture.}
    \label{fig:penalty}
\end{figure}
As shown in figure \ref{fig:sample_result2}, after we generate the result on ribbon image, just need to simply covert the ribbon image back to its original shape by reverse the method that generates the ribbon image.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.95\textwidth]{Figures/sample2_result.png}
    \caption[Desire Output on Sample Sidewalk]{Detail demonstration on desire result after applying our dynamic programming approach. Where row 1 shows the result after reshape, row 2 to 5 shows the boundaries we calculated on ribbon image from \ref{fig:Sample_Sidewalk_2}.}
    \label{fig:sample_result2}
\end{figure}

@@

We also test our approach in different cities. 
For initial trajectories, we pull the corresponding footway data from overpass-turbo \cite{overpass_turbo}.
Finding map data is challenging since there not exist much \ac{VHR} imagery from public source. 
We use the imagery data form GIS government database and treat it as input map. 
\figref{fig:oxford1} demonstrate plausible result on a sidewalks heavy area, we randomly selected a few from Oxford, OH. 
Row 1 shows that our approach are able to find the precise boundaries under trees or building shadows, with less accuracy around intersection area. 
Except above from row1, row 2 also demonstrate the camouflage with adjacent material. 
We treat the intersection part as overlapping sidewalks, and process them individually. \figref{fig:ny1} and \figref{fig:ny3} show the random samples we select from New York area. 
Predicting is heavier involved in these area since some sidewalks are partially under tree leaves instead of just shadows.

@@

% \begin{figure}
%     \centering
%     \includegraphics[width=\textwidth]{Figures/ny2.png}
%     \caption[Sample Sidewalk 5]{Demonstration on sample sidewalks with our approach. We use sidewalks that are random selected from New York area. Row 1 shows a thicker sidewalk that camouflage with adjacent road and street. Row 2 shows complex sidewalk texture with irregular shape.}
%     \label{fig:ny2}
% \end{figure}

### 2018-07-27 Chapters/5.Conclusion/Conclusion.tex
Fixing format and images

\chapter{Conclusion}
In this thesis defense, we have investigated a dynamic programming approach to segment sidewalk as a ribbon-like feature. We believe the preliminary result from our hypothesis is decent. During thesis progress, we gain the knowledge on important backgrounds and we study other feature extraction methods. Also, we develop a fully working dynamic programming algorithm with penalty control function to achieve better performance.

>>>>

\chapter{Conclusion}
In this thesis defense, a method to estimate the thickness and to refine the trajectory of a ribbon-like feature in \ac{VHR} aeral orthoimagery has been presented. 
We were able to obtain both high qualitative and quantitative results for segmenting walking paths in aerial imagery by refining \ac{OSM} features. The method is fast, and has only a few parameters with clear interpretations.
The proposed approach does not rely on extensive training on annotated data, but instead learns to fit the underlying imagery based on a trimap that is specified by parameters $\MinRadius{}, \MaxRadius{}, \MaxDistance{}$ which values that are easily determined based on the task, and are unlikely to require adjustment.

We evaluated this work by comparing it mainly to approaches for interactive image segmentation. 
Because most automatic approaches for extracting street-maps focus on extracting coarse trajectories (which we take as input) rather than on precise alignment and width estimation. 
We focused on modeling footways which are small, thin, easily occluded, and often modified due to construction.
We believe that this approach is a natural tool to improve segmentation by post-processing automatically extracted street networks, and also by enabling fast interactive annotation of aerial images by refining hand-entered paths such as those found in \ac{OSM}.

### 2018-07-27 Thesis.tex
Fixing format and images

    We present a novel approach for segmenting ribbon-like features, such as sidewalks, With the only knowledge of the approximate locations and thicknesses of thin features, this approach can register coarsely digitized footways to high-resolution aerial imagery with high accuracy. The key innovation is a dynamic-programming solution to search for the optimal set of thicknesses and offsets to an initial path. The method provides control over the rate of change of thickness and direction, and it does not rely on prior knowledge of the materials in the image. The approach will be helpful as an aid to digitization, and downstream applications such as registration of walking paths, canals, or roadways as well as for damage and change detection.

>>>>

    We present a novel approach for segmenting ribbon-like features, such as sidewalks. With the only knowledge of the approximate locations and thicknesses of thin features, this approach can register coarsely digitized footways to high-resolution aerial imagery with high accuracy. The key innovation is a dynamic-programming solution to search for the optimal set of thicknesses and offsets to an initial path. The method provides control over the rate of change of thickness and direction, and it does not rely on prior knowledge of the materials in the image. The approach will be helpful as an aid to digitization, and downstream applications such as registration of walking paths, canals, or roadways as well as for damage and change detection.
